---
title: "Data Manipulation"
author: "Eric Morris"
date: "September 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


## Import FAS csv files

Importing csv files to use as examples

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select variables

```{r}
select(litters_data, group, litter_number, gd0_weight)

select(litters_data, litter_number, gd0_weight, pups_survive)
```

... by specifying a range:

```{r}
select(litters_data, gd_of_birth:pups_survive)
select(litters_data, group, gd_of_birth:pups_survive)
```

... by saying what you want to remove:

```{r}
select(litters_data, -litter_number)
```

You can re-name variables as well using select if necessary 

```{r}
select(litters_data, group, litter_number, gest_day_0_weight = gd0_weight)
rename(litters_data, gest_day_0_weight = gd0_weight)
```

Handy helper functions:

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, litter_number, starts_with("pup"))

select(litters_data, litter_number, everything())
```

Learning assessment (select the columns containing litter number, sex, and PD ears): 

```{r}
select(pups_data, litter_number, sex, pd_ears)
```

## Filter observations

Do some filtering!

```{r}
filter(litters_data, gd0_weight < 25)
filter(litters_data, gd0_weight >= 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, !is.na(gd0_weight))

filter(litters_data, group == "Low7" | group == "Low8")

filter(litters_data, group %in% c("Low7", "Low8"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:7, !is.na(gd0_weight))
```

Learning assessment (Filter to include only pups with sex 1 and then Filter to include only pups with PD walk less than 11 and sex 2):

```{r}

filter(pups_data, sex == 1)

filter(pups_data, sex == 2, pd_walk < 11)
```

## Mutate 

Create variables

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
)
```

```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       wt_gain_squared = wt_gain^2)
```

Learning assessment (Create a variable that subtracts 7 from PD pivot and Create a variable that is the sum of all the PD variables): 

```{r}
mutate(pups_data,
       pivot_sub = pd_pivot - 7,
       pd_add = pd_ears + pd_eyes + pd_pivot + pd_walk)
```


## Arrange

Arrange the data...

```{r}
arrange(litters_data, gd0_weight)

arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

