---
title: "Tidy Data"
author: "Eric Morris"
date: "September 25, 2018"
output: github_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Gather

PULSE data

```{r}
pulse_data = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
pulse_data
```

This isn'y tidy yet...

```{r}
pulse_tidy = pulse_data %>% 
  gather(key = "visit", value = "bdi_score", bdi_score_bl:bdi_score_12m)
```

Illustrate 'separate':

```{r}
pulse_tidy %>% 
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>% 
  select(-bdi_str, -score_str) %>% 
  mutate(visit = replace(visit, visit == "bl", "00m"))
```

All together, the data import / cleaning pipeline is: 

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>% 
  gather(key = "visit", value = "bdi_score", bdi_score_bl:bdi_score_12m) %>% 
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>% 
  select(-bdi_str, -score_str) %>% 
  mutate(visit = replace(visit, visit == "bl", "00m"))
```

## Revisit FAS_litters

```{r}
litters_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
   janitor::clean_names() %>%
  separate(group, into = c("dose", "day"), 3) %>% 
  mutate(dose = tolower(dose),
         wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)
```

Learning Assessment, data cleaning/tidying

```{r}
litters_data %>% 
  select(litter_number, ends_with("weight")) %>%  
  gather(key = gd, value = weight, gd0_weight:gd18_weight) %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18)) %>% 
  arrange(litter_number)
```


## Spread

Create 'analyis_result'

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

Make it readable (spread, the reverse of gather)

```{r}
analysis_result %>% 
  spread(key = time, value = mean) %>%
  knitr::kable()
```

## Binding rows

```{r}
fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

Create final LOTR data

```{r}
lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  gather(key = sex, value = word, female:male) %>% 
  mutate(race = tolower(race))
```


## Join datasets

Load the FAS datasets and create a joined dataset via a left join

```{r}
pup_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))

FAS_data = left_join(pup_data, litter_data, by = "litter_number")
```

Learning assessment: 

```{r}
surv_os = read_csv("./data/survey_results/surv_os.csv") %>% 
  janitor::clean_names() %>% 
  rename(id = what_is_your_uni, os = what_operating_system_do_you_use)

surv_pr_git = read_csv("data/survey_results/surv_program_git.csv") %>% 
  janitor::clean_names() %>% 
  rename(id = what_is_your_uni, 
         prog = what_is_your_degree_program,
         git_exp = which_most_accurately_describes_your_experience_with_git)

left_join(surv_os, surv_pr_git)
inner_join(surv_os, surv_pr_git)
anti_join(surv_os, surv_pr_git)
anti_join(surv_pr_git, surv_os)
```

